---
title: "Fine mapping comparison"
author: "Jeremy Schwartzentruber"
date: "12/11/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)

otg = "/Users/jeremys/work/otgenetics"
```

## Selecting studies for fine-mapping comparison

We need to compare our results, done using GCTA-cojo to get independent signals + Approximate Bayes Factor (ABF) fine-mapping of the conditionally independent signals, with more modern approaches like FINEMAP. To do this we'll select a handful of GWAS to use when comparing the two methods.

First, we downloaded all studies and ordered by number of associations.

```{r}
toploci = read_tsv(file.path(otg, "gs/genetics-portal-staging/v2d/190705/toploci.tsv"))

study_locus_counts = toploci %>%
  group_by(study_id) %>%
  summarise(num_loci = n())

studies = read_tsv(file.path(otg, "gs/genetics-portal-staging/v2d/190705/studies.tsv")) %>%
  arrange(desc(num_assoc_loci)) %>%
  filter(study_id %in% study_locus_counts$study_id)

ukb_studies = studies %>%
  filter(grepl("NEALE|SAIGE", study_id))
```

Next, we selected five studies -- three quantitative traits (from Neale et al) and two binary traits (from SAIGE) for further investigation.

Here, we look at the distribution of credible set sizes (from GCTA + ABF) across loci for each of the studies.

```{r}

selected_studies = read_tsv(file.path(otg, "natgen_paper/fine_mapping/selected_ukb_studies.txt"))

finemapping = read_tsv(file.path(otg, "gs/genetics-portal-staging/v2d/190705/finemapping.tsv.gz"))
finemapping_selected = selected_studies %>%
  select(study_id, trait_reported,) %>%
  left_join(finemapping, by="study_id")
finemapping_selected_credset_sizes = finemapping_selected %>%
  group_by(study_id, trait_reported, lead_chrom, lead_pos) %>%
  summarise(credset_size = n())

ggplot(finemapping_selected_credset_sizes, aes(x=credset_size)) +
  geom_histogram() +
  facet_wrap(~trait_reported) +
  scale_x_log10()
```

